---
title: "ANNEXE Biology in philosophy of biology"
author: "Jacob Hamel-Mottiez"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
library(d3r)
library(DT)
library(forcer)
library(ggraph)
library(igraph)
library(maps)
library(plotly)
library(readxl)
library(tidyverse)
library(tidygraph)
library(viridis)
library(scales)




#dir <- "/media/louis/05E2-13B8/BioWOS/data/"
dir <- "C:/Users/jacob/OneDrive - Université Laval/DATA/"


# DATA --------------------------------------------------------------------
# on doit enlever les deux dernières lignes de revues et philo_bio parce que c'est de la bouette
revues <- read_tsv(paste0(dir,"WOS_revues.rpt")) %>% slice(1:(n()-2)) # 21 296
bio_and_philo <- read_tsv(paste0(dir,"philo_of_bio_refs.rpt")) %>% # 135 546
  slice(1:(n()-2)) %>% # 135 544
  distinct() # 67 774
liste_discipline <- read_excel(paste0(dir,"Liste_discipline.xlsx")) # 144


bio_and_philo %>% distinct() %>% nrow() #  67774
test <- bio_and_philo %>% count(OST_BK,Ordre,UID,UID_Ref,Cited_Author,Year,
                        Volume,Page,Doi,Cited_Title,Cited_Work,Patent_No,
                        Assignee,Art_No,Occurence_Order) %>% 
  filter(n==1)

test$Cited_Title
bio_and_philo %>% filter(str_detect(Cited_Title, "The importance of symbiosis in philosophy of biology"))
bio_and_philo %>% filter(str_detect(Cited_Author, "Suarez, J"))



bio_and_philo %>% filter(Cited_Work == "NULL") %>% nrow() # 22

test <- bio_and_philo %>% count(Cited_Work) %>% arrange(Cited_Work)

data_revue <- bio_and_philo %>% 
  filter(Cited_Work != "NULL") %>% 
  left_join(revues, by = c("Cited_Work" = "Revue")) %>% 
  left_join(liste_discipline, by = c("Code_discipline" = "Code_Discipline")) %>% # 67 752
  filter(!is.na(Code_discipline)) # 34 080

data_abbrev <- bio_and_philo %>% 
  filter(Cited_Work != "NULL") %>% 
  left_join(revues, by = c("Cited_Work" = "Abbrev_11")) %>% 
  left_join(liste_discipline, by = c("Code_discipline" = "Code_Discipline")) %>% 
  filter(!is.na(Code_discipline)) # 8524

data <- bind_rows(data_revue, data_abbrev)

# il y a 67344 publications sans correspondance entre Cited_Work et revues...
# ... proche de 50% ... pas très convaincant







```

## TOP CITED JOURNALS, DISCIPLINES AND SPECIALITIES
This descriptive data [to complete]

```{r TOP CITED JOURNALS, cache = TRUE}
# WHAT ARE THE TOP CITED JOURNALS IN B&P? ------------------------ ---------
top_journals <- data %>% count(Code_Revue, sort = T) %>% 
  left_join(revues)

# WHAT ARE THE TOP CITED JOURNALS, DISCIPLINES, SPECIALITIES IN B&P? ---------------------------------
top_discipline <- data %>% count(Discipline, sort = T)
top_specialite <- data %>% count(Specialite, sort = T)

top_discipline_top_journaux <- data %>% 
  filter(Code_Revue %in% top_journals$Code_Revue[1:20]) %>% 
  count(Discipline, sort = T)


fct_DT <- function(x) {datatable(x, 
         extensions = c('FixedColumns',"FixedHeader"),
          options = list(scrollX = TRUE, 
                         paging=TRUE,
                         fixedHeader=TRUE))}

fct_percent <- function(x) {
  x <-  x |>  mutate(percent = n/sum(x$n, na.rm = TRUE)*100)
}

list_table <- list(top_journals, top_discipline, top_specialite, top_discipline_top_journaux)

lapply(list_table, fct_percent)
lapply(list_table, fct_DT)


```

## PROPORTION OF DOCUMENT TYPE IN THE CORPUS
All the documents of the corpus have a Code_Dicument number. 
The variable EType_Document has more non-null entries than Type_Document variable. 

```{r DOCUMENT TYPE, echo=FALSE, cache = TRUE}
# WHAT IS THE DISTRIBUTION OF DOCUMENTS ?-----------------------------------
document_type <- read_tsv(paste0(dir, "p.ex.Liste_Document.rpt")) %>% 
  slice(1:(n()-2))
art_doc_type <- read_tsv(paste0(dir, "cited_citant_B&P.rpt")) %>% 
  slice(1:(n()-2))


document_type <- document_type |> mutate(Code_Document = as.character(Code_Document))
art_doc_type <- art_doc_type|> mutate(Code_Document = as.character(Code_Document))
art_doc_type |> filter(is.na(Code_Document) | Code_Document == "NULL") # 0

art_doc_type_count <- art_doc_type |> count(Code_Document, sort = T) 
doc_type_n <- left_join(document_type, art_doc_type_count, by = "Code_Document") |> arrange(-n)
doc_type_n <-  doc_type_n |>  mutate(percent = n/sum(doc_type_n$n, na.rm = TRUE)*100)

fct_percent(doc_type_n)
fct_DT(doc_type_n)


```

## MAPPING THE MOST PRODUCTIVE COUNTRIES

```{r MAP, cache = TRUE, echo=FALSE}
# WHAT ARE THE MOST CITED COUNTRY? ----------------------------------------
adresses <- read_tsv(paste0(dir, "articles_adress_B&P.rpt"))
art_country <- adresses |> count(Pays, sort = T) |> rename(region = Pays)

# Visualisation of it. 
world_df <-  map_data("world")
world_df$region <- toupper(world_df$region)
world_df <- left_join(world_df, art_country, by = "region")

p <- ggplot(world_df, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill= n), color = "white") +
  scale_fill_viridis(option="D", begin = 0.1, end = 0.7, direction = 1) 
ggplotly(p)
```

## NETWORK
```{r NETWORK, echo=FALSE, cache = TRUE}
# NETWORK OF CITING AND CITED ---------------------------------------------
data_network <- read_tsv(paste0(dir, "cited_citant_B&P.rpt")) |> slice(1:(n()-2))


data_network <- data_network |> select(OST_BK_Cite, OST_BK_Citant, Titre, Annee_Bibliographique)
data_network <- data_network |> rename(OST_BK = OST_BK_Cite)

# MATCHING REFS WITH SPECIALITIES
match_data <-  data |> select(OST_BK, EGrande_Discipline, EDiscipline, ESpecialite)
data_network <-  left_join(match_data, data_network, by = "OST_BK") |> filter(!is.na(OST_BK_Citant))


EDiscipline <- c("Biomedical Research", "Social Sciences", "Biology", 
                 "Humanities", "Clinical Medicine", "Psychology", 
                 "Professional Fields", "Earth and Space", 
                 "Mathematics", "Physics", "unknown")

# Vector of associated colors
edgeColor <- c("#5c7bbd", "#f08c39", "#db4743", 
            "#78c19f", "#52c497", "#a97db8", 
            "#db7f87", "#bf856c", 
            "#c9a377", "#75767d", "#50525d")

code_color <- tibble(EDiscipline, edgeColor)

data_network <- left_join(data_network, code_color, by = "EDiscipline")

          
data_network_small <- data_network |> 
  select(OST_BK, OST_BK_Citant, edgeColor) |> 
  distinct()

# Create the graph
co_citation <- data_network_small |>
  as_tbl_graph(directed = FALSE,
    nodes = "OST_BK",
    edges = "OST_Citant"
  )

num_communities <- 68  # Assuming 68 communities
hex_colors <- hue_pal()(num_communities)

nodes_names <- data_network |> select(name = OST_BK_Citant, Titre) |>
  mutate(name = as.character(name))|> 
  distinct()  
# 636 nodes names over 805, ~ 80%. 

node_weights <- co_citation |>
  activate(nodes) |>
  left_join(nodes_names, by = c("name")) |>
  mutate(name = coalesce(Titre, name)) |>
  mutate(weight = centrality_degree(mode = "in")) |> 
  filter(weight > 8) |> mutate(community = group_louvain()) |>
  mutate(color = hex_colors[as.numeric(factor(community))])
 

# VISUALISATION
g_g <- d3r::d3_igraph(node_weights)

forcer(
  data = g_g,
  forcetype = "ForceGraph2D",
  linkCurvature = 0.5, 
  NodeLabel = TRUE, 
  nodeThreeObjectExtend= FALSE,
  nodeVal = "weight", 
  #linkColor = "edgeColor",
  nodeColor = "color"
)



```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
