---
title: "REPRESENTATION OF BIOLOGY IN PHILOSOPHY OF BIOLOGY : A SYSTEMATIC BIBLIOGRAPHIC ANALYSIS"
author: "Jacob Hamel-Mottiez"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
library(d3r)
library(DT)
library(forcer)
library(ggraph)
library(igraph)
library(maps)
library(readxl)
library(tidyverse)
library(tidygraph)
library(viridis)
library(scales)
library(visNetwork)
library(ggforce)
library(concaveman)
library(plotly)
library(network)
library(particles)
library(patchwork)
library(gifski)
library(wordcloud)
library(RColorBrewer)
library(wordcloud2)

```



## Description
This projects aims at mapping the attention philosophers of biology are giving to their object : biology. Where precedent studies (e.g. Pradeu, 2017) looked at *Biology & Philosophy* compared to *PNAS*, we want to document this trend at a larger scale. 


The data is fetched through the OST database builded from Web of Science (clariviate). 


Thanks to CRSH and FRQSC who fund my research. 
Thank you François Claveau and Louis Renaud-Desjardins for your support on this project.


```{r DATA, include = FALSE, cache = TRUE}
#DIRECTORY AND DATE 
dir_od <- "C:/Users/jacob/OneDrive - Université Laval/biophilo/Data"
data <- read_csv("Data/Philosophy of Biology/PHILOSOPHY_OF_BIOLOGY_ALL_REFS.csv")
data$source_eid <- str_extract(data$source_eid, "(?<=2-s2\\.0-)\\d+")
data <- rename(data, citing_id = source_eid, cited_id = id)
```


```{r DATA, include = FALSE, cache = TRUE}
link_tbl <- data
link_tbl <- link_tbl |> select(citing_id, cited_id)
link_tbl
```

## Network construction 

```{r basic network, cache = TRUE, include = FALSE, echo=FALSE}
g <- graph_from_data_frame(link_tbl, directed = FALSE)
#cocitation(g)
```


```{r basic network, cache = TRUE, include = FALSE, echo=FALSE}
node_metrics <- g |>
  as_tbl_graph(directed = FALSE) |>
  activate(nodes) |>
  mutate(weight = centrality_degree()) |> 
  mutate(community = group_leiden()) |> 
  filter(weight > 100)

node_metrics <- node_metrics %>%
  activate(nodes) %>%
  mutate(color = factor(community)) %>%  # Assign colors based on community
  activate(edges) %>%
  mutate(color = .N()$color[from])  # Get color of source node for each edge

# Define a color palette with enough colors for all your communities
num_communities <- length(unique(node_metrics %>% activate(nodes) %>% pull(color)))
color_palette <- scales::hue_pal()(num_communities)

color_map <- tibble(
  color_palette = color_palette, 
  color = as.factor(seq(1:num_communities)),
  alpha = 0.2
)


node_metrics <- node_metrics |> 
  activate(nodes) |> 
  mutate(name, as.character(name)) |>
  left_join(color_map, by = "color") |> 
  activate(edges) |> 
  left_join(color_map, by = "color") 
```

```{r network, cache = TRUE, echo=FALSE}
link_tbl <- link_tbl |> select(citing_id, cited_id)

link_tbl_g <- graph_from_data_frame(link_tbl, directed = FALSE)

g_cocit <- bibcoupling(node_metrics) 

# I could make a function for it. 
node_metrics <- g_cocit |>
  as_tbl_graph(directed = FALSE) |>
  activate(nodes) |>
  mutate(
    #degree = centrality_degree(),
    #betweenness = centrality_betweenness(),
    #closeness = centrality_closeness()
  ) |> 
  mutate(weight = round(centrality_degree())) |> 
  mutate(community = group_leiden(objective_function = "CPM", n = 10, resolution = 0.01)) |> #objective_function = "modularity",
                                  #n = 5,
                                  #resolution = 1)) |> 
  filter(weight > 10)

# Color of the nodes and the edges
node_metrics <- node_metrics |>
  activate(nodes) |>
  mutate(color = factor(community)) |>  # Assign colors based on community
  activate(edges) |>
  mutate(color = .N()$color[from])  # Get color of source node for each edge

# Define a color palette with enough colors for all your communities
num_communities <- length(unique(node_metrics |> activate(nodes) |> pull(color)))
color_palette <- scales::hue_pal()(num_communities)

actual_communities <- sort(unique(node_metrics |> activate(nodes) |> pull(color)))

# Create the color mapping with actual community numbers
color_map <- tibble(
  color_palette = color_palette,
  color = as.factor(actual_communities),
  alpha = 0.2
)

node_metrics <- node_metrics |> 
  activate(nodes) |>
  left_join(color_map, by = "color") |> 
  activate(edges) |> 
  left_join(color_map, by = "color") 

 node_metrics |> 
  activate(nodes) |> distinct(community)

```

## 2D interactive cocitation network visualisation (VisNetwork)
```{r 2d_network , echo=FALSE}
# I need to look back at it. cocitation is a must but I need to think again for colors and everything. 

y <- visNetwork::toVisNetworkData(node_metrics)
y$nodes$value = y$nodes$weight
y$nodes$color = y$nodes$color_palette
y$edges$color = y$edges$color_palette
y$edges$width = 10
link_tbl2 <- link_tbl |> rename(label = OST_BK_Ref, time = Year) |> select(label, Cited_Author, Cited_Title, time)

y$nodes <- left_join(y$nodes, link_tbl, by = "label") |> unique()
y$nodes <- y$nodes |> mutate(label = paste0(Cited_Author,", ", time, ", ", Cited_Title)) |> filter(label != "NANA")
y$nodes$time <- lubridate::ymd(y$nodes$time, truncated = 2L)


network_2d <- visNetwork(nodes = y[[1]], edges = y[[2]], background = "#282a36") |>
  visNodes(font = list(color ="white"))|>
  visOptions(highlightNearest = TRUE, selectedBy = "community", ) |>
  visPhysics(solver = "barnesHut")
#custom_network <- visNetworkEditor(object = network_2d)
#network_2d

y$nodes |> write.csv(paste0(dir_od,"nodes_bp.csv"))
y$edges |> write.csv(paste0(dir_od,"edges_bp.csv"))

```



## 3D cocitation network visualisation 
```{r VR network , echo=FALSE}
link_tbl3 <- link_tbl |> rename(name = OST_BK_Ref) |> select(name, Cited_Author, Cited_Title, Year) |> unique()

node_metrics <- node_metrics |> 
  activate(nodes) |> left_join(link_tbl3, by = "name") 
node_metrics <- node_metrics |> activate(nodes) |> rename(name_old = name) |> mutate(name= paste0(Cited_Author, ", ", Year, ", ", Cited_Title))
node_metrics <- node_metrics |> filter(name != "NA, NA, NA")

node_metrics <- node_metrics |> activate(nodes) |> mutate(weight_small= round((weight/100)))
g_g <- d3r::d3_igraph(node_metrics)

x <- forcer(
  data = g_g,
  backgroundColor = "#282a36",
  forcetype = "ForceGraph3D",
  linkWidth = 1,
  linkCurvature = 0.5, 
  NodeLabel = TRUE, 
  nodeThreeObjectExtend= FALSE,
  nodeVal = "weight_small", 
  #nodeOpacity= "alpha",
  linkColor = "color_palette",
  nodeColor = "color_palette"
  
  #linkVisibility = 0.5,
  #linkDirectionalParticles = 1,
  #linkDirectionalParticleWidth = 5
)
x

```




## Word cloud
```{r WORD CLOUD , echo=FALSE}

# KEYWORD PLUS WORD CLOUD
keyword_count <-  ref_kw_bp %>% 
  filter(keyword != "NULL") |> 
 select(keyword) |> count(keyword, sort=TRUE)

wordcloud(words = keyword_count$keyword, freq = keyword_count$n, min.freq = 1, max.words=100, random.order=FALSE, rot.per=0.35, colors=brewer.pal(8, "Dark2"))
wordcloud2(keyword_count)


# KEYWORD PLUS WORD CLOUD
keyword_plus_count <-  ref_kw_bp %>% 
  filter(keyword_plus != "NULL") |> 
 select(keyword_plus) |> count(keyword_plus, sort=TRUE)


wordcloud(words = keyword_plus_count$keyword_plus, freq = keyword_plus_count$n, min.freq = 1, max.words=100, random.order=FALSE, rot.per=0.35, colors=brewer.pal(8, "Dark2"))
wordcloud2(keyword_plus_count)

```


```{r, cache = TRUE}

link_tbl_g <- graph_from_data_frame(link_tbl, directed = FALSE)
weight_filter <-  800

g <- cocitation(link_tbl_g) |> as_tbl_graph(directed = FALSE)

# I could make a function for it. 
node_metrics <- g |>
  as_tbl_graph(directed = FALSE) |>
  activate(nodes) |>
  mutate(
    #degree = centrality_degree(),
    #betweenness = centrality_betweenness(),
    #closeness = centrality_closeness()
  ) |> 
  mutate(weight = round(centrality_degree())) |> 
  mutate(community = group_louvain()) |> 
  filter(weight > weight_filter)

# Color of the nodes and the edges
node_metrics <- node_metrics |>
  activate(nodes) |>
  mutate(color = factor(community)) |>  # Assign colors based on community
  activate(edges) |>
  mutate(color = .N()$color[from])  # Get color of source node for each edge

# Define a color palette with enough colors for all your communities
num_communities <- length(unique(node_metrics |> activate(nodes) |> pull(color)))
color_palette <- scales::hue_pal()(num_communities)

actual_communities <- sort(unique(node_metrics |> activate(nodes) |> pull(color)))

# Create the color mapping with actual community numbers
color_map <- tibble(
  color_palette = color_palette,
  color = as.factor(actual_communities),
  alpha = 0.2
)

node_metrics <- node_metrics |> 
  activate(nodes) |>
  left_join(color_map, by = "color") |> 
  activate(edges) |> 
  left_join(color_map, by = "color") 

```

## 2D interactive cocitation network visualisation (VisNetwork)
```{r 2d_network , echo=FALSE}
# I need to look back at it. cocitation is a must but I need to think again for colors and everything. 

y <- visNetwork::toVisNetworkData(node_metrics)
y$nodes$value = y$nodes$weight
y$nodes$color = y$nodes$color_palette
y$edges$color = y$edges$color_palette
y$edges$width = 10
link_tbl$id = link_tbl$scopus_id

y$nodes <- left_join(y$nodes, link_tbl, by = "id") |> unique()
y$nodes <- y$nodes |> mutate(label = paste0(author,", ", year, ", ", ifelse(is.na(title), sourcetitle, title))) |> filter(label != "NA, NA, NA")
#y$nodes$time <- lubridate::ymd(y$nodes$time, truncated = 2L)

y$nodes |> write.csv(paste0(dir_od,"nodes_biol_th.csv"))
y$edges |> write.csv(paste0(dir_od,"edges_biol_th.csv"))

```


Get top references by cluster
```{r, eval = F}
top_ref_com <- y$nodes |> filter(community == seq(1, 10)) |>  slice_max(order_by = weight, n = 6, by = color) 

#p <- ggplot(top_ref_com, aes(x = reorder(label, weight), y = weight, fill = color)) + 
geom_bar(stat = "identity", show.legend = F) + 
  facet_wrap(~color, scales = "free")  + 
  coord_flip()+
  theme_void()

#ggplotly(p)

```



