---
title: "REPRESENTATION OF BIOLOGY IN PHILOSOPHY OF BIOLOGY : A SYSTEMATIC BIBLIOGRAPHIC ANALYSIS"
author: "Jacob Hamel-Mottiez"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
library(d3r)
library(DT)
library(forcer)
library(ggraph)
library(igraph)
library(maps)
library(readxl)
library(tidyverse)
library(tidygraph)
library(viridis)
library(scales)
library(visNetwork)
library(ggforce)
library(concaveman)
library(plotly)
library(network)
library(particles)
library(patchwork)
library(gifski)
library(wordcloud)
library(RColorBrewer)
library(wordcloud2)

```



## Description
This projects aims at mapping the attention philosophers of biology are giving to their object : biology. Where precedent studies (e.g. Pradeu, 2017) looked at *Biology & Philosophy* compared to *PNAS*, we want to document this trend at a larger scale. 


The data is fetched through the OST database builded from Web of Science (clariviate). 


Thanks to CRSH and FRQSC who fund my research. 
Thank you François Claveau and Louis Renaud-Desjardins for your support on this project.


```{r DATA, include = FALSE, cache = TRUE}
#DIRECTORY AND DATE 
dir_od <- "C:/Users/jacob/OneDrive - Université Laval/DATA/"
date <- "2024-09-23"
video_path <- "C:/Users/jacob/Videos/Screen Recordings/Screen Recording 2024-10-02 124946.mp4"

# COLUMN NAMES FOR FOLLOWING DATA
ref_col <- c(
  "OST_BK",
  "UID",
  "OST_BK_Ref",
  "UID_Ref",
  "Cited_Author",
  "Year",
  "Cited_Title",
  "Cited_Work"
)

art_col <- c(
  "OST_BK",
  "citing_year",
  "UID",
  "Titre",
  "Nb_Reference",
  "Type_Document", 
  "Revue",
  "Abrev",
  "Abstract"
)

ref_auth_col <-  c(
  "OST_BK_Ref",
  "UID_Ref",
  "first_name",
  "last_name",
  "seq_no"
)

ref_abstr_col <- c(
  "OST_BK_Ref", 
  "UID_Ref", 
  "Abstract"
)
  
art_auth_col <- c(
  "OST_BK", 
  "first_name", 
  "last_name",
  "seq_no"
)

ref_cit_top_col <- c(
  "OST_BK_Ref", 
  "UID_Ref", 
  "citation_topic"
)

ref_kw_col <- c(
  "OST_BK_Ref", 
  "UID_Ref", 
  "keyword",
  "keyword_plus"
)


# DATA
bp_db <- read_csv(paste0(dir_od,"bp_article_db.csv"), skip = 1) # This is the data from B&P and not WoS. 
ref_bp <- read_delim(paste0(dir_od,
                            "bp_references_2024-09-23.csv"), delim = ";", col_names = ref_col)
art_bp <- read_delim(paste0(dir_od, 
                            "bp_articles_alt_2024-09-28.csv"),  delim = ";", col_names = art_col)
ref_auth_bp <- read_delim(paste0(dir_od,
                                 "bp_references-authors_2024-09-23.csv"), delim = ";", col_names = ref_auth_col)
art_auth_bp <- read_delim(paste0(dir_od, 
                                 "bp_articles-authors_2024-09-23.csv"),  delim = ";", col_names = art_auth_col)
ref_abstr_bp <- read_delim(paste0(dir_od,
                                  "bp_references-abstract_2024-09-23.csv"), delim = ";", col_names= ref_abstr_col)


ref_cit_top_bp <- read_delim(paste0(dir_od,
                                  "bp_references-citations-topic_2024-10_02.csv"), delim = ";", col_names = ref_cit_top_col) |> print(n = 100)

art_adresse_bp <- read_delim(paste0(dir_od,
                                  "bp_articles_adresses_2024-10-02.csv"), delim = ";") # Weird, why no columns need to be specify?
art_org_bp <- read_delim(paste0(dir_od,
                                  "bp_articles-organizations_2024-10-02.csv"), delim = ";") # Weird, why no columns need to be specify?
ref_kw_bp <- read_delim(paste0(dir_od,
                                  "bp_references-keywords_2024-10_02.csv"), delim = ";", col_names =  ref_kw_col)


link_tbl <- ref_bp |> select(OST_BK, OST_BK_Ref, Cited_Author, Cited_Title,Year) |> filter(OST_BK_Ref != "NULL")


```


## Descriptive data 
```{r AUTHORS, echo=FALSE, cache = TRUE, warning= FALSE}
# WHAT IS THE DISTRIBUTION OF THE CITING DOCUMENT? 
top_document_cited <-  art_bp |>
  count(Type_Document, sort = T)

# WHO ARE THE MOST CITING AUTHORS? 
top_citing_author <- art_auth_bp |>
  filter(first_name != "NULL",last_name != "NULL") |>
  count(first_name, last_name, sort = T) |> unique()

# WHO ARE THE MOST CITED AUTHORS? 
top_cited_author <- ref_auth_bp |>
  filter(first_name != "NULL",last_name != "NULL") |>
  count(first_name, last_name, sort = T) |> unique()
## NOT very useful the most cited authors. 


## HERE, I CAN DO NOTHING WITH THE DATA I CURRENTLY HAVE.

# # WHAT ARE THE TOP CITED JOURNALS IN B&P? ------------------------ ---------
# top_journals <- data %>% count(Code_Revue, sort = T) %>% 
#   left_join(revues)
# 
# # WHAT ARE THE TOP CITED JOURNALS, DISCIPLINES, SPECIALITIES IN B&P? ---------------------------------
# top_discipline <- data %>% count(Discipline, sort = T)
# top_specialite <- data %>% count(Specialite, sort = T)
# 
# top_discipline_top_journaux <- data %>% 
#   filter(Code_Revue %in% top_journals$Code_Revue[1:20]) %>% 
#   count(Discipline, sort = T)

##  END. 
rows_displayed = 10

fct_percent <- function(x) {
  x <-  x |>  mutate(percent = n/sum(x$n, na.rm = TRUE)*100) |>
    mutate(across(percent, round, 3)) |> 
    slice_max(n, n = rows_displayed)
}


fct_DT <- function(x, title) {
        dt <- datatable(x, 
          caption = title, 
          extensions = c('FixedColumns',"FixedHeader"),
          options = list(scrollX = TRUE, 
                         paging=TRUE,
                         fixedHeader=TRUE))
}



list_table <- list(top_document_cited, top_citing_author, top_cited_author) 

percent_tbl <- lapply(list_table, fct_percent)


x <- fct_DT(x = percent_tbl[[1]], title = "Types of documents in B&P")
x
y <- print(fct_DT(x = percent_tbl[[2]], title = "Most prominent author published in B&P")) 
y
z <- print(fct_DT(x = percent_tbl[[3]], title = "Most cited author in B&P article's references")) 
z
# 

```

```{r WOS VS SPRINGER, echo=FALSE}

# Nombre d'articles de la revue étudiée ----------------------------------
art_bp %>% distinct(OST_BK) %>% nrow() # 1673


art_year_WoS <- art_bp %>% select(OST_BK, citing_year) %>% 
  filter(citing_year < 2023) |> # data from 2023 is not complete. 
  distinct() %>% 
  count(citing_year) |> 
  rename(N = n)

art_year_WoS <- art_year_WoS |> mutate(FROM = "WoS")


art_year_db <- bp_db %>% select(YEAR, ARTICLES) %>% 
  filter(YEAR < 2023) |>
  distinct() 

art_year_db <- art_year_db |> 
  rename(citing_year = YEAR, N = ARTICLES) |> 
  mutate(FROM = "B&P SRINGER")


full_table <- rbind(art_year_db, art_year_WoS) 

# Plot
color_map <- c("#5ae284", "#4e5c68")

plot_art <- ggplot(full_table, aes(x = citing_year, y = N, fill = FROM)) +
  geom_col(position = "dodge") + 
  ggtitle("Articles répertoriés par WoS (OST) vs réperoriés sur Springer pour B&P") +
  scale_fill_manual(values = color_map)
plot_art
```


## Network construction 

```{r basic network, cache = TRUE, include = FALSE, echo=FALSE}
g <- graph_from_data_frame(link_tbl, directed = FALSE)
cocitation(g)


node_metrics <- g |>
  as_tbl_graph(directed = FALSE) |>
  activate(nodes) |>
  mutate(weight = centrality_degree()) |> 
  mutate(community = group_leiden()) |> 
  filter(weight > 30)

node_metrics <- node_metrics %>%
  activate(nodes) %>%
  mutate(color = factor(community)) %>%  # Assign colors based on community
  activate(edges) %>%
  mutate(color = .N()$color[from])  # Get color of source node for each edge

# Define a color palette with enough colors for all your communities
num_communities <- length(unique(node_metrics %>% activate(nodes) %>% pull(color)))
color_palette <- scales::hue_pal()(num_communities)

color_map <- tibble(
  color_palette = color_palette, 
  color = as.factor(seq(1:num_communities)),
  alpha = 0.2
)


node_metrics <- node_metrics |> 
  activate(nodes) |> 
  mutate(name, as.character(name)) |>
  left_join(color_map, by = "color") |> 
  activate(edges) |> 
  left_join(color_map, by = "color") 
```  

```{r basic network visualisation, echo=FALSE}
y <- visNetwork::toVisNetworkData(node_metrics)
y$nodes$value = y$nodes$weight
y$nodes$color = y$nodes$color_palette
y$edges$color = y$edges$color_palette
y$edges$width = 10

link_tbl2 <- link_tbl |> rename(label = OST_BK_Ref) |> select(label, Cited_Author, Cited_Title)

y$nodes <- left_join(y$nodes, link_tbl2, by = "label") |> unique()

y$nodes <- y$nodes |> mutate(label = paste0(Cited_Author, Cited_Title)) |> filter(label != "NA, NA, NA")

simple_network <- visNetwork(nodes = y[[1]], edges = y[[2]], background = "#282a36") |>
  visNodes(font = list(color ="white"))|>
  visOptions(highlightNearest = TRUE, selectedBy = "community", )

```

```{r network, cache = TRUE, echo=FALSE}

link_tbl_g <- graph_from_data_frame(link_tbl, directed = FALSE)
weight_filter <-  100

g <- cocitation(link_tbl_g) |> as_tbl_graph(directed = FALSE)

# I could make a function for it. 
node_metrics <- g |>
  as_tbl_graph(directed = FALSE) |>
  activate(nodes) |>
  mutate(
    #degree = centrality_degree(),
    #betweenness = centrality_betweenness(),
    #closeness = centrality_closeness()
  ) |> 
  mutate(weight = round(centrality_degree())) |> 
  mutate(community = group_leiden(objective_function = "CPM", n = 10, resolution = 0.01)) |> #objective_function = "modularity",
                                  #n = 5,
                                  #resolution = 1)) |> 
  filter(weight > weight_filter)

# Color of the nodes and the edges
node_metrics <- node_metrics |>
  activate(nodes) |>
  mutate(color = factor(community)) |>  # Assign colors based on community
  activate(edges) |>
  mutate(color = .N()$color[from])  # Get color of source node for each edge

# Define a color palette with enough colors for all your communities
num_communities <- length(unique(node_metrics |> activate(nodes) |> pull(color)))
color_palette <- scales::hue_pal()(num_communities)

actual_communities <- sort(unique(node_metrics |> activate(nodes) |> pull(color)))

# Create the color mapping with actual community numbers
color_map <- tibble(
  color_palette = color_palette,
  color = as.factor(actual_communities),
  alpha = 0.2
)

node_metrics <- node_metrics |> 
  activate(nodes) |>
  left_join(color_map, by = "color") |> 
  activate(edges) |> 
  left_join(color_map, by = "color") 

 node_metrics |> 
  activate(nodes) |> distinct(community)

```

## 2D interactive cocitation network visualisation (VisNetwork)
```{r 2d_network , echo=FALSE}
# I need to look back at it. cocitation is a must but I need to think again for colors and everything. 

y <- visNetwork::toVisNetworkData(node_metrics)
y$nodes$value = y$nodes$weight
y$nodes$color = y$nodes$color_palette
y$edges$color = y$edges$color_palette
y$edges$width = 10

link_tbl2 <- link_tbl |> rename(label = OST_BK_Ref, time = Year) |> select(label, Cited_Author, Cited_Title, time)

y$nodes <- left_join(y$nodes, link_tbl2, by = "label") |> unique()
y$nodes <- y$nodes |> mutate(label = paste0(Cited_Author,", ", time, ", ", Cited_Title)) |> filter(label != "NA, NA, NA")
y$nodes$time <- lubridate::ymd(y$nodes$time, truncated = 2L)


network_2d <- visNetwork(nodes = y[[1]], edges = y[[2]], background = "#282a36") |>
  visNodes(font = list(color ="white"))|>
  visOptions(highlightNearest = TRUE, selectedBy = "community", ) |>
  visPhysics(solver = "barnesHut")
#custom_network <- visNetworkEditor(object = network_2d)
#network_2d

y$nodes |> write.csv(paste0(dir_od,"nodes_bp.csv"))
y$edges |> write.csv(paste0(dir_od,"edges_bp.csv"))


```

Get top references by cluster
```{r}
nodes <- read_csv(paste0(dir_od,"nodes_bp.csv"))
edges <-  read_csv(paste0(dir_od,"edges_bp.csv"))




top_ref_com <- nodes |> slice_max(order_by = weight, n = 8, by = community) 

p <- top_ref_com %>%
  mutate(label = fct_reorder(label, weight)) %>%
  ggplot(aes(x = label, y = weight, fill = color)) +
    geom_col(show.legend = FALSE) +
    scale_fill_identity() +  # Use color values from the color column directly
    theme_void() +
    coord_flip() +
    facet_wrap(~community, scales = "free")
p

ggplotly(p)


```

```{r}

```




## 3D cocitation network visualisation 
```{r VR network , echo=FALSE}
link_tbl3 <- link_tbl |> rename(name = OST_BK_Ref) |> select(name, Cited_Author, Cited_Title, Year) |> unique()

node_metrics <- node_metrics |> 
  activate(nodes) |> left_join(link_tbl3, by = "name") 
node_metrics <- node_metrics |> activate(nodes) |> rename(name_old = name) |> mutate(name= paste0(Cited_Author, ", ", Year, ", ", Cited_Title))
node_metrics <- node_metrics |> filter(name != "NA, NA, NA")

node_metrics <- node_metrics |> activate(nodes) |> mutate(weight_small= round((weight/100)))
g_g <- d3r::d3_igraph(node_metrics)

x <- forcer(
  data = g_g,
  backgroundColor = "#282a36",
  forcetype = "ForceGraph3D",
  linkWidth = 1,
  linkCurvature = 0.5, 
  NodeLabel = TRUE, 
  nodeThreeObjectExtend= FALSE,
  nodeVal = "weight_small", 
  #nodeOpacity= "alpha",
  linkColor = "color_palette",
  nodeColor = "color_palette"
  
  #linkVisibility = 0.5,
  #linkDirectionalParticles = 1,
  #linkDirectionalParticleWidth = 5
)
x

```




## Word cloud
```{r WORD CLOUD , echo=FALSE}

# KEYWORD PLUS WORD CLOUD
keyword_count <-  ref_kw_bp %>% 
  filter(keyword != "NULL") |> 
 select(keyword) |> count(keyword, sort=TRUE)

wordcloud(words = keyword_count$keyword, freq = keyword_count$n, min.freq = 1, max.words=100, random.order=FALSE, rot.per=0.35, colors=brewer.pal(8, "Dark2"))
wordcloud2(keyword_count)


# KEYWORD PLUS WORD CLOUD
keyword_plus_count <-  ref_kw_bp %>% 
  filter(keyword_plus != "NULL") |> 
 select(keyword_plus) |> count(keyword_plus, sort=TRUE)


wordcloud(words = keyword_plus_count$keyword_plus, freq = keyword_plus_count$n, min.freq = 1, max.words=100, random.order=FALSE, rot.per=0.35, colors=brewer.pal(8, "Dark2"))
wordcloud2(keyword_plus_count)

```


```{r, cache = TRUE}

link_tbl_g <- graph_from_data_frame(link_tbl, directed = FALSE)
weight_filter <-  800

g <- cocitation(link_tbl_g) |> as_tbl_graph(directed = FALSE)

# I could make a function for it. 
node_metrics <- g |>
  as_tbl_graph(directed = FALSE) |>
  activate(nodes) |>
  mutate(
    #degree = centrality_degree(),
    #betweenness = centrality_betweenness(),
    #closeness = centrality_closeness()
  ) |> 
  mutate(weight = round(centrality_degree())) |> 
  mutate(community = group_louvain()) |> 
  filter(weight > weight_filter)

# Color of the nodes and the edges
node_metrics <- node_metrics |>
  activate(nodes) |>
  mutate(color = factor(community)) |>  # Assign colors based on community
  activate(edges) |>
  mutate(color = .N()$color[from])  # Get color of source node for each edge

# Define a color palette with enough colors for all your communities
num_communities <- length(unique(node_metrics |> activate(nodes) |> pull(color)))
color_palette <- scales::hue_pal()(num_communities)

actual_communities <- sort(unique(node_metrics |> activate(nodes) |> pull(color)))

# Create the color mapping with actual community numbers
color_map <- tibble(
  color_palette = color_palette,
  color = as.factor(actual_communities),
  alpha = 0.2
)

node_metrics <- node_metrics |> 
  activate(nodes) |>
  left_join(color_map, by = "color") |> 
  activate(edges) |> 
  left_join(color_map, by = "color") 

```

## 2D interactive cocitation network visualisation (VisNetwork)
```{r 2d_network , echo=FALSE}
# I need to look back at it. cocitation is a must but I need to think again for colors and everything. 

y <- visNetwork::toVisNetworkData(node_metrics)
y$nodes$value = y$nodes$weight
y$nodes$color = y$nodes$color_palette
y$edges$color = y$edges$color_palette
y$edges$width = 10
link_tbl$id = link_tbl$scopus_id

y$nodes <- left_join(y$nodes, link_tbl, by = "id") |> unique()
y$nodes <- y$nodes |> mutate(label = paste0(author,", ", year, ", ", ifelse(is.na(title), sourcetitle, title))) |> filter(label != "NA, NA, NA")
#y$nodes$time <- lubridate::ymd(y$nodes$time, truncated = 2L)

y$nodes |> write.csv(paste0(dir_od,"nodes_biol_th.csv"))
y$edges |> write.csv(paste0(dir_od,"edges_biol_th.csv"))

```


Get top references by cluster
```{r, eval = F}
top_ref_com <- y$nodes |> filter(community == seq(1, 10)) |>  slice_max(order_by = weight, n = 6, by = color) 

#p <- ggplot(top_ref_com, aes(x = reorder(label, weight), y = weight, fill = color)) + 
geom_bar(stat = "identity", show.legend = F) + 
  facet_wrap(~color, scales = "free")  + 
  coord_flip()+
  theme_void()

#ggplotly(p)

```



