---
title: "References to biology and to philosophy of biology"
author: "Jacob Hamel-Mottiez"
date: "`r Sys.Date()`"
output:
  html_document:
    #runtime: shiny
    toc: true
    toc_float:
      collapsed: true
    code_folding: hide
    
---

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(plotly)
library(viridis)
library(DT) # for clean and interactive datatable in .rmd. 
library(data.table)
library(knitr)
library(readxl)
```

## Loading the data

First, we need to load the data. As one can see, we here load generalist
philosophy of science journals. Our 4 initial journals of philosophy of
biology include :

-   BIOLOGY AND PHILOSOPHY,

-   BIOLOGICAL THEORY,

-   STUDIES IN HISTORY AND PHILOSOPHY OF SCIENCE PART C,

-   HISTORY AND PHILOSOPHY OF THE LIFE SCIENCES.

Our generalists philosophy of science journals include :

-   PHILOSOPHY_OF_SCIENCE,

-   BRITISH JOURNAL FOR THE PHILOSOPHY OF SCIENCE,

-   SYNTHESE,

-   ERKENNTNIS,

-   EUROPEAN JOURNAL FOR PHILOSOPHY OF SCIENCE,

-   INTERNATIONAL STUDIES IN THE PHILOSOPHY OF SCIENCE,

-   JOURNAL FOR GENERAL PHILOSOPHY OF SCIENCE,

-   FOUNDATIONS OF SCIENCE

Moreover, after looking through PhilPapers selection of philosophy of
biology edited by Manolo Martinez, we added the following journals :

-   ACTA BIOTHEORETICA,

-   BEHAVIORAL AND BRAIN SCIENCES,

-   BIOESSAYS,

-   BIOSEMIOTICS.


```{r DATA, message = FALSE, warning = FALSE, echo = FALSE, result = 'asis', cache = TRUE}
dir_data <- "C:/Users/jacob/OneDrive - Université Laval/biophilo/Data//pybiblio/GENERAL PHILOSOPHY OF SCIENCE/"

PHILOSOPHY_OF_BIOLOGY_ALL_4_INITIAL_JOURNALS <- read_csv("C:/Users/jacob/OneDrive - Université Laval/biophilo/Data//pybiblio/SPECIALIZED PHILOSOPHY OF BIOLOGY/ARTICLES_SPECIAL_PHILO_BIO.csv")

for_name_general_philo_of_science = c(
  "PHILOSOPHY_OF_SCIENCE",
  "BRITISH_JOURNAL_FOR_THE_PHILOSOPHY_OF_SCIENCE", 
  #"SYNTHESE", 
  "ERKENNTNIS", 
  "EUROPEAN_JOURNAL_FOR_PHILOSOPHY_OF_SCIENCE", 
  "INTERNATIONAL_STUDIES_IN_THE_PHILOSOPHY_OF_SCIENCE", 
  "JOURNAL_FOR_GENERAL_PHILOSOPHY_OF_SCIENCE", 
  "FOUNDATIONS_OF_SCIENCE", 
  #"ACTA_BIOTHEORETICA", 
  #"BEHAVIORAL_AND_BRAIN_SCIENCES",
  #"BIOESSAYS",
  #"BIOSEMIOTICS",
  "STUDIES_IN_HISTORY_AND_PHILOSOPHY_OF_SCIENCE"
  #"BEHAVIORAL_AND_BRAIN_SCIENCES"
)

# Create an empty list to store the dataframes
articles_data <- list()
references_data <- list()

# Load each CSV file and assign it to the list with its clean name
for (i in seq_along(for_name_general_philo_of_science)) {
  articles_file_path <- paste0(dir_data, for_name_general_philo_of_science[i], ".csv")
  articles_data[[for_name_general_philo_of_science[i]]] <- read_csv(articles_file_path)
  
  references_file_path <- paste0(dir_data, for_name_general_philo_of_science[i], "_refs_pyblio.csv")
  references_data[[paste0(for_name_general_philo_of_science, "_refs")[i]]] <- read_csv(references_file_path)
}

# Attach the dataframes to the global environment with their clean names
invisible(list2env(articles_data, envir = .GlobalEnv))
invisible(list2env(references_data, envir = .GlobalEnv))

```

```{r PROCESSING FUNCTIONS}
# CLEAN ARTICLES AND REFERENCES -------------------------------------
clean_references_fct <- function(references) {  
  references$sourcetitle <- toupper(references$sourcetitle)
  references$citing_id <- str_extract(references$citing_eid, "(?<=2-s2\\.0-)\\d+")
  references <- rename(references, cited_id = id, cited_title = title)
  references$citing_eid <- NULL
  references <- rename(references, cited_year = coverDate)
  return(references)
}

clean_articles_fct <- function(articles){
  articles$citing_id <- str_extract(articles$eid, "(?<=2-s2\\.0-)\\d+")
  articles$eid <- NULL
  articles <- rename(articles, citing_year = coverDate,
                     citing_category = subtypeDescription,
                     citing_title = title)
  return(articles)
}  

# MERGE ARTICLES AND THEIR REFERENCES -------------------------------------
merge_fct <- function(articles, references) { 
  # Merging
  reference_is_article <- left_join(
    references,
    articles |> select(citing_id, citing_category, citing_year, publicationName, citing_title), 
    by = "citing_id") |> 
    filter(citing_category == "Article")
  return(reference_is_article)
}



# DETERMINE WHICH REFERENCES ARE FROM OUR INITIAL CORPUS ------------------
is_philo_bio_fct <- function(reference_is_article){
  # is the references from our corpus? TRUE or FALSE
  distinct_journals <- PHILOSOPHY_OF_BIOLOGY_ALL_4_INITIAL_JOURNALS |> distinct(publicationName) 
  distinct_journals$publicationName <- toupper(distinct_journals$publicationName)
  
  # Get the mean by years for 1) total refs, 2) refs from philo. of bio. and 3) refs not from philo. and bio
  reference_is_article <- reference_is_article |>
    mutate(is_philo_bio_journal = sourcetitle %in% distinct_journals$publicationName) |>
    group_by(citing_id) |>
    mutate(
      refs_from_philo_bio = sum(is_philo_bio_journal, na.rm = TRUE),
      refs_not_from_philo_bio = sum(!is_philo_bio_journal, na.rm = TRUE),
      total_refs = n()
    ) |>
    ungroup()
  return(reference_is_article)
}


# CALCULATE THE PROPORTION OF REFERENCES TO PHILO. OF BIO./TOTAL R --------

proportion_fct <- function(reference_is_article) {
  proportion_table <- reference_is_article |> select(citing_id, citing_title, publicationName, total_refs, refs_from_philo_bio, refs_not_from_philo_bio) |> arrange(desc(refs_from_philo_bio)) |> distinct()
  proportion_table <- proportion_table |> mutate(ratio_philo_bio = refs_from_philo_bio/total_refs*100)
  return(proportion_table)
}



fct_percent <- function(x) {
  dt <-  x |>  mutate(percent = n/sum(x$n, na.rm = TRUE)*100) |>
    mutate(across(percent, round, 3)) 
  dt
}

fct_DT <- function(x) {
        dt <- DT:: datatable(head(x, 1000),
          options = list(scrollX = TRUE, 
                         paging=TRUE,
                         pageLength = 5))
        dt
}

```

```{r, message = FALSE, warning = FALSE, echo = FALSE, result = 'asis', cache = TRUE}
# Assuming your vector of data frame names is called df_names
 # example names

# Define a function to process each data frame
process_df <- function(for_name_general_philo_of_science) {
  # Clean the articles and references
  df_cleaned <- clean_articles_fct(get(for_name_general_philo_of_science))
  df_refs_cleaned <- clean_references_fct(get(paste0(for_name_general_philo_of_science, "_refs")))

  # Merge and classify
  reference_is_article <- merge_fct(df_cleaned, df_refs_cleaned)
  reference_is_article <- is_philo_bio_fct(reference_is_article)

  # Calculate proportions
  proportion_tbl <- proportion_fct(reference_is_article)
  
  return(proportion_tbl)
}

# Apply the function to each data frame
proportions_list <- lapply(for_name_general_philo_of_science, process_df)


```

```{r FUNCTION FOR VIZ, message = FALSE, warning = FALSE, echo = FALSE, result = 'asis', cache = TRUE }
year_fct <- function(reference_is_article) {
  philo_bio_articles_in_general <- reference_is_article
  
  # Ensure citing_year is numeric
  philo_bio_articles_in_general <- philo_bio_articles_in_general %>%
    mutate(citing_year = as.numeric(format(citing_year, "%Y")))
  
  # Calculate total references per citing_id first
  philo_bio_articles_in_general <- philo_bio_articles_in_general %>%
    group_by(citing_id) %>%
    mutate(total_refs = n()) %>%
    ungroup()
  
  # Group by citing_year and summarize the mean counts
  philo_bio_articles_in_general <- philo_bio_articles_in_general %>%
    group_by(citing_year) %>%
    summarise(
      mean_n_ref_total = mean(total_refs, na.rm = TRUE),
      mean_n_ref_not_bio_philo = mean(refs_not_from_philo_bio, na.rm = TRUE),
      mean_n_ref_bio_philo = mean(refs_from_philo_bio, na.rm = TRUE)
    ) %>%
    ungroup()
  
  return(philo_bio_articles_in_general)
}

```

# Mean number of references to philosophy of biology articles in philosophy articles published in generalist philosophy of science journals

```{r VIZ RATIO DISTRIBUTION, message = FALSE, warning = FALSE, echo = FALSE, result = 'asis', cache = TRUE}
process_and_plot <- function(for_name_general_philo_of_science) {
  # Assuming `reference_is_article` is already cleaned and classified
  df_cleaned <- clean_articles_fct(get(for_name_general_philo_of_science))
  df_refs_cleaned <- clean_references_fct(get(paste0(for_name_general_philo_of_science, "_refs")))

  # Merge and classify
  reference_is_article <- merge_fct(df_cleaned, df_refs_cleaned)
  reference_is_article <- is_philo_bio_fct(reference_is_article)

  # Extract year data and select relevant columns
  philo_bio_articles_in_general <- year_fct(reference_is_article)
  philo_bio_articles_in_general <- philo_bio_articles_in_general |> select(citing_year, 
                                                                          mean_n_ref_total, 
                                                                          mean_n_ref_not_bio_philo, 
                                                                          mean_n_ref_bio_philo)

  # Reshape data to long format
  data_long <- philo_bio_articles_in_general |> 
    pivot_longer(cols = c(mean_n_ref_total, mean_n_ref_bio_philo),
                 names_to = "category", values_to = "count") |> 
    distinct()

  # Create the plot
  g <- ggplot(data_long, aes(x = citing_year, y = count, fill = category)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(
      title = paste0(for_name_general_philo_of_science),
      x = "Year", 
      y = "Mean N. of References", 
      fill = NULL) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),   # Bold, centered title
      axis.text.x = element_text(size=14, face = "bold"),              # Bold x-axis text
      axis.text.y = element_text(size=14, face = "bold"),               # Bold y-axis text
      axis.title.x = element_text(size=14, face="bold"),    
      axis.title.y = element_text(size=14, face="bold")
    )
  
  return(g)
}

# Apply the function to each data frame
plots_list <- lapply(for_name_general_philo_of_science, process_and_plot)
plots_list
```

# Density of the ratio references to specialized philosophy of biology over the remaining references. 

```{r, message = FALSE, warning = FALSE, cache = TRUE}
density_and_plot <- function(for_name_general_philo_of_science, color) {
  df_cleaned <- clean_articles_fct(get(for_name_general_philo_of_science))
  df_refs_cleaned <- clean_references_fct(get(paste0(for_name_general_philo_of_science, "_refs")))
  
  reference_is_article <- merge_fct(df_cleaned, df_refs_cleaned)
  reference_is_article <- is_philo_bio_fct(reference_is_article)

  philo_bio_articles_in_general <- year_fct(reference_is_article)
  philo_bio_articles_in_general <- philo_bio_articles_in_general |> 
    select(citing_year, mean_n_ref_total, mean_n_ref_not_bio_philo, mean_n_ref_bio_philo)

  proportion_tbl <- proportion_fct(reference_is_article)
  
  v2 <- ggplot(proportion_tbl, aes(x = ratio_philo_bio)) +
    geom_density(fill = color, color = color, alpha = 0.5) + 
    labs(title = for_name_general_philo_of_science) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      axis.text.x = element_text(size = 14, face = "bold"),
      axis.text.y = element_text(size = 14, face = "bold"),
      axis.title.x = element_text(size = 14, face = "bold"),
      axis.title.y = element_text(size = 14, face = "bold")
    )

  v2
}
```


```{r, message = FALSE, warning = FALSE, cache = TRUE}
# Generate a vector of distinct colors from the Viridis palette
num_plots <- length(for_name_general_philo_of_science)
viridis_colors <- viridis(num_plots, option = "D")  # "D" is a good general-purpose Viridis option

# Apply the function with different colors
density_list <- mapply(density_and_plot, for_name_general_philo_of_science, viridis_colors, SIMPLIFY = FALSE)

# Check the plots
density_list

```

```{r, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
combined_data <- lapply(for_name_general_philo_of_science, function(name) {
  df_cleaned <- clean_articles_fct(get(name))
  df_refs_cleaned <- clean_references_fct(get(paste0(name, "_refs")))

  reference_is_article <- merge_fct(df_cleaned, df_refs_cleaned)
  reference_is_article <- is_philo_bio_fct(reference_is_article)
  
  proportion_tbl <- proportion_fct(reference_is_article)
  proportion_tbl$dataset <- name  # Add dataset name as a column
  return(proportion_tbl)
}) |> bind_rows()


```

# When we merge all the general philosophy of science journals togheter

```{r, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
combined_data_philo_bio <- combined_data |> filter(ratio_philo_bio >= 0.5) 
# 2,822 out of 33, 061 articles. So a bit less than 10%. 

combined_density <- combined_data |> ggplot(aes(x = ratio_philo_bio)) +
  geom_density(color = "red", fill = "red", alpha = 0.3)

ggplotly(combined_density)


```

# Corpus of philosophy of biology
Now, we bind the general philosophy of science that is "hidden"
philosophy of biology, with our corpus of specialized philosophy of
biology. We used a ratio of references to specialized philosophy of
biology versus to something else of 0.5. In plain english, it means that
each articles from general philosophy of science journals need to have
at least one out of two references from our 4 initial specialized
journals in philosophy of biology. Obviously, we put this number so
highly because we want to limit false positive. We are confident that
with this method we detect praticaly only philosophy of biology
articles.

All the references to those articles will be considered as a citation *to philosophy of biology*. 

```{r, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
combined_data <- lapply(for_name_general_philo_of_science, function(name) {
  df_cleaned <- clean_articles_fct(get(name))
  df_cleaned  <-  df_cleaned |> mutate(across(everything(), as.character)) 
  return(df_cleaned)
}) |> bind_rows()

article_philo_bio_general <- combined_data[combined_data$citing_id %in% combined_data_philo_bio$citing_id,] 
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
PHILOSOPHY_OF_BIOLOGY_ALL_4_INITIAL_JOURNALS  <-  PHILOSOPHY_OF_BIOLOGY_ALL_4_INITIAL_JOURNALS |> mutate(across(everything(), as.character)) 
full_data_philo_bio <- bind_rows(article_philo_bio_general, PHILOSOPHY_OF_BIOLOGY_ALL_4_INITIAL_JOURNALS)
```

We now have a corpus of 4,287 articles of philosophy of biology from
specialized journals and 2, 732 articles from the most important
journals in general philosophy of science, accounting for a total of 7,
109 articles. We can now proceed to identify what kind of biology they
cite. To do this, we first look at what are the main journals they are
citing. We then manually annotate each one of them to keep only the most
important biology journals.

```{r, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
final_table <- bind_rows(references_data)
final_table <- clean_references_fct(final_table)
```


```{r, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
final_table <- final_table[final_table$citing_id %in% full_data_philo_bio$citing_id,] 
#write_csv(final_table, paste0(dir_od, "final_table.csv"))
```

We get around 400, 000 references, which is a lot. Now, lets look at the
main journals. Without surprise, we find generalist philosophy of
science journals such as the *BJPS* and *Studies in history and
Philosophy of Biological and Biomedical Sciences*. Without surprise, we
find generalist science journals such as *Nature* and *Science*. We find
also generalist journals of philosophy such as *Journal of Philosophy*.
However, what interst us are the specialized journals in biology. We see
some of them such as EVOLUTION, CELL, NATURE REVIEWS GENETICS, GENETICS,
BEHAVIORAL, CURRENT BIOLOGY, PLOS BIOLOGY, SYSTEMATIC BIOLOGY, ECOLOGY,
etc.

Before going further, it is important to note that here, there are
differences in sourcetitle and that they may not reflect the correct
journal. For instance, BRITISH JOURNAL FOR THE PHILOSOPHY OF SCIENCE is
valid but not THE BRITISH JOURNAL FOR THE PHILOSOPHY OF SCIENCE when
querying Scopus. We thus pass our data into Openrefine to get cleaner
data.

```{r, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
# Some functions to display nice data table and to add percentages automatically
final_table_OR <-  read_csv("C:/Users/jacob/OneDrive - Université Laval/biophilo/Data/openrefine/most-cited-journal-in-philo-bio_openrefine.csv")
cited_journals <- final_table_OR |> select(cited_id, sourcetitle, cited_title) |> filter(!is.na(cited_title)) |> count(sourcetitle)  |> arrange(desc(n))

cited_journals <- fct_percent(cited_journals)
fct_DT(cited_journals)

```

Once this has been done, it gives us our first idea of what philosophy
of biology cites. For computational reasons, it is not currently
possible to fetch all the journals in biology (Careful, there is about 70 entries that are not found in Scopus, I need to manually check them).

We thus proceeded to select the most important. To do so, we first
looked at their citescore, a metric that calculate the ratio of citation
over the number of publication. Scopus API gives us access to this
metric for each year of the journals. Hence, we calculate the citescore
mean and multiply it by the number of references it gets from philosophy
of biology.

We ended up selecting 60 journals with the highest weighted citescore : 

-   JOURNAL OF THEORETICAL BIOLOGY\
-   EVOLUTION\
-   CELL\
-   PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA\
-   NATURE REVIEWS GENETICS\
-   GENETICS\
-   AMERICAN NATURALIST\
-   CURRENT BIOLOGY\
-   TRENDS IN COGNITIVE SCIENCES\
-   TRENDS IN ECOLOGY AND EVOLUTION\
-   JOURNAL OF EVOLUTIONARY BIOLOGY\
-   QUARTERLY REVIEW OF BIOLOGY\
-   BIOSCIENCE\
-   COGNITION\
-   PHILOSOPHICAL TRANSACTIONS OF THE ROYAL SOCIETY B BIOLOGICAL SCIENCES\
-   NATURE GENETICS\
-   NEURON\
-   PLOS BIOLOGY\
-   SYSTEMATIC BIOLOGY\
-   ECOLOGY\
-   NATURE NEUROSCIENCE\
-   ANNALS OF THE NEW YORK ACADEMY OF SCIENCES\
-   JOURNAL OF MOLECULAR BIOLOGY\
-   NATURE REVIEWS MICROBIOLOGY\
-   NATURE REVIEWS NEUROSCIENCE\
-   BIOLOGICAL REVIEWS\
-   TRENDS IN MICROBIOLOGY\
-   JOURNAL OF NEUROSCIENCE\
-   MOLECULAR BIOLOGY AND EVOLUTION\
-   TRENDS IN GENETICS\
-   NATURE REVIEWS MOLECULAR CELL BIOLOGY\
-   GENOME BIOLOGY\
-   GENOME RESEARCH\
-   ANNUAL REVIEW OF MICROBIOLOGY\
-   ANNUAL REVIEW OF NEUROSCIENCE\
-   ECOLOGY LETTERS\
-   ANNUAL REVIEW OF ECOLOGY EVOLUTION AND SYSTEMATICS\
-   ANNUAL REVIEW OF GENETICS\
-   NATURE REVIEWS CANCER\
-   AMERICAN JOURNAL OF HUMAN GENETICS\
-   TRENDS IN NEUROSCIENCES\
-   MICROBIOLOGY AND MOLECULAR BIOLOGY REVIEWS\
-   TRENDS IN BIOCHEMICAL SCIENCES\

Hence, all the articles published in those journals will be considered biology articles. Conversely, all the references one of those articles, in philosophy of biology, will be considered as a citation *to biology*. 


# Example with TRENDS IN BIOCHEMICAL SCIENCES and MICROBIOLOGY AND MOLECULAR BIOLOGY REVIEWS
The basic idea here is to first take the references from philosophy of biology and from biology journals. We label references that are from one of the biology journals we previously identified as a citation *to biology*. The rest is considered reference *not_to_biology*.  Second, we compute the number of times biology references' are cited in philosophy and in biology (we also compute their rank).

```{r, message = FALSE, warning = FALSE, cache = TRUE}
# Your list of biology journals
bio_journals <- c(
  "JOURNAL OF THEORETICAL BIOLOGY", "EVOLUTION", "CELL",
  "PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA",
  "NATURE REVIEWS GENETICS", "GENETICS", "AMERICAN NATURALIST",
  "CURRENT BIOLOGY", "TRENDS IN COGNITIVE SCIENCES",
  "TRENDS IN ECOLOGY AND EVOLUTION", "JOURNAL OF EVOLUTIONARY BIOLOGY",
  "QUARTERLY REVIEW OF BIOLOGY", "BIOSCIENCE", "COGNITION",
  "PHILOSOPHICAL TRANSACTIONS OF THE ROYAL SOCIETY B: BIOLOGICAL SCIENCES",
  "NATURE GENETICS", "NEURON", "PLOS BIOLOGY", "SYSTEMATIC BIOLOGY",
  "ECOLOGY", "NATURE NEUROSCIENCE", "ANNALS OF THE NEW YORK ACADEMY OF SCIENCES",
  "JOURNAL OF MOLECULAR BIOLOGY", "NATURE REVIEWS MICROBIOLOGY",
  "NATURE REVIEWS NEUROSCIENCE", "BIOLOGICAL REVIEWS",
  "TRENDS IN MICROBIOLOGY", "JOURNAL OF NEUROSCIENCE",
  "MOLECULAR BIOLOGY AND EVOLUTION", "TRENDS IN GENETICS",
  "NATURE REVIEWS MOLECULAR CELL BIOLOGY", "GENOME BIOLOGY",
  "GENOME RESEARCH", "ANNUAL REVIEW OF MICROBIOLOGY",
  "ANNUAL REVIEW OF NEUROSCIENCE", "ECOLOGY LETTERS",
  "ANNUAL REVIEW OF ECOLOGY, EVOLUTION, AND SYSTEMATICS",
  "ANNUAL REVIEW OF GENETICS", "NATURE REVIEWS CANCER",
  "AMERICAN JOURNAL OF HUMAN GENETICS", "TRENDS IN NEUROSCIENCES",
  "MICROBIOLOGY AND MOLECULAR BIOLOGY REVIEWS", "TRENDS IN BIOCHEMICAL SCIENCES"
)
```

```{r, message = FALSE, warning = FALSE, cache = TRUE}
# REFERENCES CORPUS ------------------------------------------------------------------
# Biology
dir_data <- "C:/Users/jacob/OneDrive - Université Laval/biophilo/Data/pybiblio/"

TRENDS_IN_BIOCHEMICAL_SCIENCES_refs_pyblio <- read_csv(paste0(dir_data, "TRENDS_IN_BIOCHEMICAL_SCIENCES_refs_pyblio.csv"))
MICROBIOLOGY_AND_MOLECULAR_BIOLOGY_REVIEWS_refs_pyblio <- read_csv(paste0(dir_data,"MICROBIOLOGY_AND_MOLECULAR_BIOLOGY_REVIEWS_refs_pyblio.csv"))
JOURNAL_OF_THEORETICAL_BIOLOGY_refs_pyblio <- read_csv(paste0(dir_data, "JOURNAL_OF_THEORETICAL_BIOLOGY_refs_pyblio.csv"))

# Philosophy of Biology 
PHILOSOPHY_OF_BIOLOGY_ALL_REFS <- read_csv("C:/Users/jacob/OneDrive - Université Laval/biophilo/Data/Philosophy of Biology/PHILOSOPHY_OF_BIOLOGY_ALL_REFS.csv")
```


```{r, message = FALSE, warning = FALSE, cache = TRUE}
# WHAT IS MOST CITED IN PHILOSOPHY AND BIOLOGY ----------------------------
tbs_counts <- rbind(TRENDS_IN_BIOCHEMICAL_SCIENCES_refs_pyblio, MICROBIOLOGY_AND_MOLECULAR_BIOLOGY_REVIEWS_refs_pyblio, JOURNAL_OF_THEORETICAL_BIOLOGY_refs_pyblio) |> 
  count(id, name = "citations_in_biology")

pob_counts <- PHILOSOPHY_OF_BIOLOGY_ALL_REFS |> 
  count(id, name = "citations_in_philosophy")

# Full join datasets and get the rank of each references relative to where it is cited. 
full_refs <- tbs_counts |> 
  full_join(pob_counts, by = "id") |> 
  mutate(
    rank_biology = dense_rank(desc(citations_in_biology)),  # Rank in biology
    rank_philosophy = dense_rank(desc(citations_in_philosophy))  # Rank in philosophy
  )
```


```{r, message = FALSE, warning = FALSE, cache = TRUE}
# WHICH REFERENCES ARE FROM BIOLOGY? ---------------------------------------
# Some formating first
PHILOSOPHY_OF_BIOLOGY_ALL_REFS <-  rename(PHILOSOPHY_OF_BIOLOGY_ALL_REFS, citing_eid = source_eid)
PHILOSOPHY_OF_BIOLOGY_ALL_REFS$...1 = NULL

all_references_info <- rbind(TRENDS_IN_BIOCHEMICAL_SCIENCES_refs_pyblio, MICROBIOLOGY_AND_MOLECULAR_BIOLOGY_REVIEWS_refs_pyblio,  PHILOSOPHY_OF_BIOLOGY_ALL_REFS)
all_references_info |> distinct()


test <- left_join(full_refs, all_references_info |> 
            select(title,sourcetitle,publicationyear, id), by = "id") |>
  arrange(-desc(rank_biology)) |> distinct()
```


```{r, message = FALSE, warning = FALSE, cache = TRUE}
# Cleaning
most_common <- function(x) {
  if (all(is.na(x))) return(NA)  # Handle all NA cases
  x[which.max(tabulate(match(x, unique(x))))]  # Find most frequent value
}

# Cleaning function
clean_references <- function(df) {
  df %>%
    group_by(id) %>%
    mutate(
      title = most_common(title),
      sourcetitle = most_common(sourcetitle),
      publicationyear = most_common(coverDate)
    ) %>%
    ungroup()
}

# Apply function to your dataset
cleaned_data <- clean_references(test) |> distinct()
```


```{r, message = FALSE, warning = FALSE, cache = TRUE}
# Add a new column classifying references
cleaned_data$sourcetitle <- toupper(cleaned_data$sourcetitle)
cleaned_data <- cleaned_data %>%
  mutate(
    ref_category = ifelse(sourcetitle %in% bio_journals, "ref_biology", "ref_not_from_our_biology_corpus"))


# HOW MANY REFERENCES TO BIOLOGY ARTICLES ARE PRESENT IN BOTH CORPUS?  ------------------------
results <- cleaned_data |> #filter(ref_category == "ref_biology") |> 
  filter(!is.na(rank_biology & rank_philosophy)) 

fct_DT(results)
```






