---
title: "Fetching data"
author: "Jacob Hamel-Mottiez"
date: "`r Sys.Date()`"
output: html_document
---
This file document how we fetched the data for Web of Science and Scopus via SQL query for the first and via API through RStudio for the second. 


# Web of Science B&P 
## DATA STRUCTURE 
We are working with Web of Science (clarivate) database via Albator. The
access is provided by Vincent Larivière.

All the information linking cited and citing documents is established via a key called "OST_BK" which is an unique identifier. 

There are two ways ways in which you can deal with the data. 
  1) Take the .dbo tables which are from WoS ;  
  2) Take the .pex tables which are from OST. 

What should influence your choice is the following methodological considerations : 
The table that links the cited documents with the citing documents is constructed as follow :   
  1) The cited document must be present in the WOS (Web of Science) corpus.
  2) The cited document must be an article, a note, or a review (with a `code_document` less than 4).
  3) The documents citing or cited must have been published between 1900 and the present day.
  4) The documents citing or cited must have at least one author listed in the `summary_name` table.

In short, if you want all the documents cited for a given citing journal, take .dbo. If you only want the scientific articles, then take .pex because it is cleaner. 


## COMPLETNESS OF THE DATA
When looking at B&P (Code_Revue = 2229), we were able to fetch a total of 67 774 cited documents. 
When we looked at which got a Cited_Title, it dropped to 43 828.
When we looked at which had a Cited_Title and an OST_BK, it dropped to 36 335 (~ 31 436 lost in total).

```{sql SCRIPT, eval=FALSE}
/****** Fetching relevant corpora
Project: Biology in philosophy
created by Jacob Hamel-Mottiez and Louis Renaud-Desjarding on 2024-09-23. 
Database : Web of Science

// SETUP //
We run many tests on Albator SQL Server. The cleanest data output is when you execute 
the query, right-click on the result and save it as a .csv file. 

You can easily save on onedrive and fetch the data elsewhere

// DATA STRUCTURE //
We are working with Web of Science (clarivate) database via Albator. The
access is provided by Vincent Lariviere.

All the information linking cited and citing documents is established via a key called "OST_BK" which is an unique identifier. 

There are two ways ways in which you can deal with the data. 
  1) Take the .dbo tables which are from WoS ;  
  2) Take the .pex tables which are from OST. 

What should influence your choice is the following methodological considerations : 
The table that links the cited documents with the citing documents is constructed as follow :   
  1) The cited document must be present in the WOS (Web of Science) corpus.
  2) The cited document must be an article, a note, or a review (with a `code_document` less than 4).
  3) The documents citing or cited must have been published between 1900 and the present day.
  4) The documents citing or cited must have at least one author listed in the `summary_name` table.

In short, if you want all the documents cited for a given citing journal, take .dbo. If you only want the scientific articles, then take .pex because it is cleaner. 


// COMPLETNESS OF THE DATA //
When looking at B&P (Code_Revue = 2229), we were able to fetch a total of 67 774 cited documents. 
When we looked at which got a Cited_Title, it dropped to 43 828.
When we looked at which had a Cited_Title and an OST_BK, it dropped to 36 335 (~ 31 436 lost in total).  


******/

-- alternative table (to verify with Louis) 
-- The idea was to select only relevant column in art. and to add Etype doc. 
SELECT art.OST_BK
  , art.Annee_Bibliographique as citing_year
  , art.UID
  , art.Titre
  , art.Nb_Reference
  , doc.EType_Document as Type_Document
  , rev.Revue
  , rev.Abbrev_11 as Abrev
  , abst.Abstract as Abstract
 FROM [WoS].[pex].[Article] as art
  LEFT JOIN [WoS].[pex].[Liste_revue] as rev 
  on art.Code_Revue = rev.Code_Revue
  LEFT JOIN [WoS].[dbo].[Abstract] as abst
  on art.OST_BK = abst.OST_BK
  LEFT JOIN [WoS].[pex].[Liste_Document] as doc
  on art.Code_Document = doc.Code_Document
  WHERE art.Code_Revue = 2229

-- Citing documents (articles published in B&P)
SELECT art.*
  ,Revue
  ,Abbrev_11
  ,abst.Abstract
  FROM [WoS].[pex].[Article] as art
  LEFT JOIN [WoS].[pex].[Liste_revue] as rev
  on art.Code_Revue = rev.Code_Revue
  LEFT JOIN [WoS].[dbo].[Abstract] as abst
  on art.OST_BK = abst.OST_BK
  WHERE art.Code_Revue = 2229

-- Author citing (the authors' of citing documents)
SELECT art.OST_BK
	, aut.First_Name
	, aut.Last_Name
	, aut.Seq_No
  FROM [WoS].[pex].[Article] as art
  LEFT JOIN [WoS].[dbo].[Summary_Name] as aut
  on art.OST_BK = aut.OST_BK
  WHERE art.Code_Revue = 2229

-- Cited documents (references of citing documents)
SELECT ref.[OST_BK]
      ,ref.[UID]
	  ,id.[OST_BK] as OST_BK_Ref
      ,ref.[UID_Ref]
      ,ref.[Cited_Author]
      ,ref.[Year]
      ,ref.[Cited_Title]
      ,ref.[Cited_Work]
  FROM [WoS].[dbo].[Reference] as ref
  LEFT JOIN [WoS].[pex].[Article] as art
  ON ref.OST_BK = art.OST_BK
  LEFT JOIN [WoS].[dbo].[Dictionnaire_ID] as id
  ON ref.UID_Ref = id.UID
  WHERE art.Code_Revue = 2229 
  
  
--  AND ref.[Cited_Work] IS NOT NULL
--  #67 752

--	AND id.[OST_BK] IS NOT NULL 
--  # 43 828 out of 67 774.

--	AND ref.[Cited_Title] IS NOT NULL 
--  # 36 335  out of 67 774.

/* MEANING OF THE COMMENTED CONDITIONS
Those two conditions enable us to look at how many references 
don't have an OST_BK (and thus no information about cited documents) 
and which cited documents had no title. 
*/

-- Cited abstract (references' abstract)
SELECT DISTINCT id.[OST_BK] as OST_BK_Ref
      ,ref.[UID_Ref]
	    ,abst.Abstract
  FROM [WoS].[dbo].[Reference] as ref
  LEFT JOIN [WoS].[pex].[Article] as art
  ON ref.OST_BK = art.OST_BK
  LEFT JOIN [WoS].[dbo].[Dictionnaire_ID] as id
  ON ref.UID_Ref = id.UID
  LEFT JOIN [WoS].[dbo].[Abstract] as abst
  on id.OST_BK = abst.OST_BK
  WHERE art.Code_Revue = 2229
	AND abst.Abstract IS NOT NULL 

-- alt. cited abstract (references' abstract) (with keywords and Keywords plus)
SELECT DISTINCT id.[OST_BK] as OST_BK_Ref
      ,ref.[UID_Ref]
	    ,abst.[Abstract]
		,keyw.[Keyword] 
		,keywP.[Keyword] as KeywordP
  FROM [WoS].[dbo].[Reference] as ref
  LEFT JOIN [WoS].[pex].[Article] as art
  ON ref.OST_BK = art.OST_BK
  LEFT JOIN [WoS].[dbo].[Dictionnaire_ID] as id
  ON ref.UID_Ref = id.UID
  LEFT JOIN [WoS].[dbo].[Abstract] as abst
  on id.OST_BK = abst.OST_BK
  LEFT JOIN [WoS].[dbo].[Keyword] as keyw
  on id.OST_BK = keyw.OST_BK
  LEFT JOIN [WoS].[dbo].[Keyword_Plus] as keywP
  on id.OST_BK = keywP.OST_BK
  WHERE art.Code_Revue = 2229
	AND abst.Abstract IS NOT NULL 
  
-- Cited authors (references' authors)
SELECT DISTINCT id.[OST_BK] as OST_BK_Ref
      ,ref.[UID_Ref]
	  ,aut.First_Name
	  ,aut.Last_Name
	  ,aut.Seq_No
  FROM [WoS].[dbo].[Reference] as ref
  LEFT JOIN [WoS].[pex].[Article] as art
  ON ref.OST_BK = art.OST_BK
  LEFT JOIN [WoS].[dbo].[Dictionnaire_ID] as id
  ON ref.UID_Ref = id.UID
  LEFT JOIN [WoS].[dbo].[Summary_Name] as aut
  on id.OST_BK = aut.OST_BK
  WHERE art.Code_Revue = 2229
	AND aut.Seq_No IS NOT NULL

-- The addresses of citing documents
SELECT contrib.*
  FROM [WoS].[dbo].Address as contrib
  LEFT JOIN [WoS].[pex].[Article] as art
  ON contrib.OST_BK = art.OST_BK
  WHERE art.Code_Revue = 2229

-- The organizations of citing documents
SELECT contrib.*
  FROM [WoS].[dbo].Address_Organization as contrib
  LEFT JOIN [WoS].[pex].[Article] as art
  ON contrib.OST_BK = art.OST_BK
  WHERE art.Code_Revue = 2229
```


## Scopus B&P
Now, we turn to Scopus. We must thank Aurélien Goutsmedt which made us aware of an API given by Scopus. It eased out our work substantially (for more information, see his blog [here](https://aurelien-goutsmedt.com/post/extracting-biblio-data-1/)).

```{r, eval = FALSE}
# API -----------------------------------------------------------
api_key <- "db914ab9d5f084e447fa55aa8c441393"
set_api_key(api_key)

insttoken <- "d7e05f3ac9d63a6d51b9170681078e96"
insttoken <- inst_token_header(insttoken)

```


```{r, eval=FALSE}
# BIOLOGY & PHILOSOPHY
bio_philo_query <- rscopus::scopus_search("EXACTSRCTITLE(Biology and Philosophy)", 
                                     view = "COMPLETE",
                                     headers = insttoken)

bio_philo_raw <- gen_entries_to_df(bio_philo_query$entries)
```

```{r, eval=FALSE}
# DIVISION OF INFORMATION INTO DIFFERENT TABLES.  ------------------------
bio_philo_papers <- bio_philo_raw$df |> clean_names()# Papers
bio_philo_affiliations <- bio_philo_raw$affiliation |> clean_names()# Affiliations
bio_philo_authors <- bio_philo_raw$author |> clean_names()# Authors

write_csv(bio_philo_papers, paste0(dir_od, "bio_philo_papers.csv"))
write_csv(bio_philo_affiliations, paste0(dir_od, "bio_philo_affiliations.csv"))
write_csv(bio_philo_authors, paste0(dir_od, "bio_philo_authors.csv"))

```

```{r, eval=FALSE}
## This chunk of code below is to catch all the references. When there is more than 40 references in an article, Scopus API is not able to get them all. Hence we need a loop to collect them all. 
citing_articles <- bio_philo_papers$dc_identifier # extracting the IDs of our articles
citation_list <- list()

# Loop to collect all the references.
for(i in 1:length(citing_articles)){
  citations_query <- abstract_retrieval(citing_articles[i],
                                        identifier = "scopus_id",
                                        view = "REF",
                                        headers = insttoken)
  if(!is.null(citations_query$content$`abstracts-retrieval-response`)){ # Checking if the article has some references before collecting them

    nb_ref <- as.numeric(citations_query$content$`abstracts-retrieval-response`$references$`@total-references`)
    citations <- gen_entries_to_df(citations_query$content$`abstracts-retrieval-response`$references$reference)$df

    if(nb_ref > 40){ # The loop to collect all the references
      nb_query_left_to_do <- floor((nb_ref) / 40)
      cat("Number of requests left to do :", nb_query_left_to_do, "\n")
      for (j in 1:nb_query_left_to_do){
        cat("Request n°", j , "\n")
        citations_query <- abstract_retrieval(citing_articles[i],
                                              identifier = "scopus_id",
                                              view = "REF",
                                              startref = 40*j+1,
                                              headers = insttoken)
        citations_sup <- gen_entries_to_df(citations_query$content$`abstracts-retrieval-response`$references$reference)$df
        citations <- bind_rows(citations, citations_sup)
      }
    }

    citations <- citations |>
      as_tibble(.name_repair = "unique") |>
      select_if(~!all(is.na(.)))

    citation_list[[citing_articles[i]]] <- citations
  }
}


bio_philo_references <- bind_rows(citation_list, .id = "citing_art") # References
write_csv(bio_philo_references, paste0(dir_od, "bio_philo_references.csv"))

```

```{r API QUERY, eval=FALSE}
# BIOLOGICAL THEORY QUERY
bio_th_query <- rscopus::scopus_search("EXACTSRCTITLE(\"Biological Theory\"", 
                                     view = "COMPLETE",
                                     headers = insttoken)

bio_th_raw <- gen_entries_to_df(bio_th_query$entries)
```

```{r, eval=FALSE}
# DIVISION OF INFORMATION INTO DIFFERENT TABLES
bio_th_papers <- bio_th_raw$df |> clean_names() # Papers
bio_th_affiliations <- bio_th_raw$affiliation |> clean_names() # Affiliations
bio_th_authors <- bio_th_raw$author |> clean_names() # Authors

write_csv(bio_th_papers, paste0(dir_od, "bio_th_papers.csv"))
write_csv(bio_th_affiliations, paste0(dir_od, "bio_th_affiliations.csv"))
write_csv(bio_th_authors, paste0(dir_od, "bio_th_authors.csv"))

```

```{r, eval=FALSE}
## This chunk of code below is to catch all the references. When there is more than 40 references in an article, Scopus API is not able to get them all. Hence we need a loop to collect them all. 

citing_articles <- bio_th_papers$dc_identifier # extracting the IDs of our articles
citation_list <- list()

# Loop to collect all the references.
for(i in 1:length(citing_articles)){
  citations_query <- abstract_retrieval(citing_articles[i],
                                        identifier = "scopus_id",
                                        view = "REF",
                                        headers = insttoken)
  if(!is.null(citations_query$content$`abstracts-retrieval-response`)){ # Checking if the article has some references before collecting them

    nb_ref <- as.numeric(citations_query$content$`abstracts-retrieval-response`$references$`@total-references`)
    citations <- gen_entries_to_df(citations_query$content$`abstracts-retrieval-response`$references$reference)$df

    if(nb_ref > 40){ # The loop to collect all the references
      nb_query_left_to_do <- floor((nb_ref) / 40)
      cat("Number of requests left to do :", nb_query_left_to_do, "\n")
      for (j in 1:nb_query_left_to_do){
        cat("Request n°", j , "\n")
        citations_query <- abstract_retrieval(citing_articles[i],
                                              identifier = "scopus_id",
                                              view = "REF",
                                              startref = 40*j+1,
                                              headers = insttoken)
        citations_sup <- gen_entries_to_df(citations_query$content$`abstracts-retrieval-response`$references$reference)$df
        citations <- bind_rows(citations, citations_sup)
      }
    }

    citations <- citations |>
      as_tibble(.name_repair = "unique") |>
      select_if(~!all(is.na(.)))

    citation_list[[citing_articles[i]]] <- citations
  }
}

bio_th_references <- bind_rows(citation_list, .id = "citing_art") # References
write_csv(bio_th_references, paste0(dir_od, "bio_th_references.csv"))

```